---
title: "dplyr sample codes"
author: "Sedi Abbasnasab"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
require(dplyr)
```
#load data
```{r}
demo <- read.csv("demographics.csv")
View(demo)
```
#filtering
#owners with the income greater than 55
```{r}
demo2 <- filter(demo, income>55)
View(demo2)
```
#owners with income equal to 60
```{r}
demo2 <- filter(demo, income==60)
View(demo2)
```
#married owners
```{r}
demo2 <- filter(demo, marital=='Married')
View(demo2)
```
#married owners with the income greater than 55
```{r}
demo2 <- filter(demo, marital=='Married', income>55)
View(demo2)
```
#multiple conditions
```{r}
demo2 <- filter(demo, marital=='Married', income>55, carcat=='Luxury')
View(demo2)
```
#owners with age between 35 and 45
```{r}
demo2 <- filter(demo, age>=35, age<=45)
View(demo2)
```
#owners who are married OR the income greater than 55
```{r}
demo2 <- filter(demo, income>55 | marital=='Married')
View(demo2)
```
#owners aged 42, 45 and 60
```{r}
demo2 <- filter(demo, age %in% c(42, 45, 60))
View(demo2)
```
#owners with the ages different from 42, 45 and 60
```{r}
demo2 <- filter(demo, !age %in% c(42, 45, 60))
View(demo2)
```

#Select
#select the variables age, marital status and gender
```{r}
demo2 <- select(demo, age, marital, gender)
View(demo2)
```
#use colon to select a range of variables
```{r}
demo2 <- select(demo, income:carcat)
View(demo2)
```
#select the third and the fifth variable
```{r}
demo2 <- select(demo, c(3, 5))
View(demo2)
```
#select the variables from the second to the forth
```{r}
demo2 <- select(demo, 2:4)
View(demo2)
```
#remove variables age and income
```{r}
demo2 <- select(demo, -age, -income)
View(demo2)
```
#remove the sixth and the eighth variable
```{r}
demo2 <- select(demo, -6, -8)
View(demo2)
```
#remove the first three variables
```{r}
demo2 <- select(demo, -1:3)
View(demo2)
```
#select the variables that contain a certain string
```{r}
demo2 <- select(demo, contains("ret"))
View(demo2)
```
#select the variables that start with the specified string
```{r}
demo2 <- select(demo, starts_with("in"))
```
#select the variables that end with the specified string
```{r}
demo2 <- select(demo, ends_with("cat"))
View(demo2)
```

#Mutate
#add a new variable containing the difference between income and car price
```{r}
demo2 <- mutate(demo, income2 = income * 10)
View(demo2)
```
#add another variable containing the difference between income and car price
```{r}
demo2 <- mutate(demo, diff = income - carpr)
View(demo2)
```
#Creat a vector and add it to data frame as a variable
#the vector lenght must be equal to the number of rows in the data frame
```{r}
y <- rnorm(510, 0, 1)
demo2 <- mutate(demo2, new_var=y)
View(demo2)
```
#divide variables 1 and 5 to 1000
```{r}
demo2 <- mutate(demo2, across(c(1,5), ~ ./1000, .names = "{col}_2"))
View(demo2)
```
#devide all the numeric variables to 1000
```{r}
demo2 <- mutate(demo, across(where(is.numeric), ~ ./1000, .names = "{col}_2"))
View(demo2)
```
#multiply colums 1 and 5 by 3
```{r}
demo2 <- mutate(demo, across(c(1,5), ~ .*3, .names = "{col}_new"))
View(demo2)
```
#compute the means of the columns 1 and 5
```{r}
demo2 <- mutate(demo, across(c(1,5), .fns = mean, .names = "{col}_mean"))
View(demo2)
```
#compute the means of all numeric variables
```{r}
demo2 <- mutate(demo, across(where(is.numeric), .fns = mean, .names = "{col}_mean"))
View(demo2)
```
#use the filter() function inside the mutate() function
#divide columns 1,3 and 5 by 1000 for the retired owners only
```{r}
demo2 <- mutate(filter(demo, retired =="Yes"), across(c(1,3, 5), ~ ./1000, .names = "{col}_2"))
View(demo2)
```

#Arrange
#sort by income in ascending order
```{r}
demo2 <- arrange(demo, income)
View(demo2)
```
#sort by income in descending order
```{r}
demo2 <- arrange(demo, desc(income))
View(demo2)
```
#sort by education level from A to Z
```{r}
demo2 <- arrange(demo, educ)
View(demo2)
```
#sort by education level from Z to A
```{r}
demo2 <- arrange(demo, desc(educ))
View(demo2)
```
#sort by income and education, in ascending order
```{r}
demo2 <- arrange(demo, income, educ)
View(demo2)
```
#sort by education and income in ascending order
```{r}
demo2 <- arrange(demo, educ, income)
View(demo2)
```
#sort by income in ascending order and by education in descending order
```{r}
demo2 <- arrange(demo, income, desc(educ))
View(demo2)
```

#Summarise
#compute the mean of income
```{r}
meaninc <- summarise(demo, avginc=mean(income, na.rm=TRUE))
View(meaninc)
typeof(meaninc)
```
#compute the standard deviation of income
```{r}
sdinc <- summarise(demo, sdincome=sd(income, na.rm = TRUE))
View(sdinc)
```
#compute the sum of the income values
```{r}
suminc <- summarise(demo, suminc=sum(income, na.rm = TRUE))
View(suminc)
```
#compute the median of the income
```{r}
medinc <- summarise(demo, medinc=median(income, na.rm = TRUE))
View(medinc)
```
#find the maximum income
```{r}
maxinc <- summarise(demo, maxinc=max(income, na.rm = TRUE))
View(maxinc)
```
#compute the variance of income
```{r}
varinc <- summarise(demo, varinc=var(income, na.rm = TRUE))
View(varinc)
```
#the number of rows in the dataset
```{r}
summarise(demo, n())
```
#compute several statistics at once
```{r}
summarise(demo, avginc=mean(income, na.rm = TRUE), sdinc=sd(income, na.rm = TRUE), varinc=var(income, na.rm = TRUE))
```
#combine summarise with other functions, like filter
#compute the average income, the standard deviation of the income, and the number of rows for the observations where the age is greater than 35
```{r}
summarise(filter(demo, age>35), avginc=mean(income, na.rm = TRUE), sdinc=sd(income, na.rm = TRUE), n())
```
#compute the average income, the standard deviation of the income and the number of rows for the male owners
```{r}
summarise(filter(demo, gender=='Male'), avginc=mean(income, na.rm = TRUE), sdinc=sd(income, na.rm = TRUE), n())
```
#compute the mean for the variables age and car price
```{r}
summarise(demo, across(c(1,5), mean))
```
#compute the mean for all numeric variables
```{r}
summarise(demo, across(where(is.numeric), mean))
```
#compute the mean for all the numeric variables but for the luxury car owners only
```{r}
summarise(filter(demo, carcat=='Luxury'), across(where(is.numeric), mean))
```
#compute the standard deviation for all the numeric variables but for the luxury car owners only
```{r}
summarise(filter(demo, carcat=='Luxury'), across(where(is.numeric), sd))
```
#comupte the mean for the variables that start with the letter "c"
```{r}
summarise(demo, across(starts_with("c"), mean))
```
#compute the mean for the numeric variables that start with the letter "c"
```{r}
summarise(demo, across(starts_with("c") & where(is.numeric), mean))
```
#extract the variables in a seprate data frame and use everything argument in the across function
```{r}
demo2 <- select(demo, age, income, carpr)
View(demo2)
summarise(demo2, across(everything(), mean))
```

# group by
# group the demo data frame by the categories of the variable education
```{r}
demo2 <- group_by(demo, educ)
View(demo2)
```
# compute the average income
```{r}
summarise(demo, avginc=mean(income))
```
# compute the standard deviation of the variable car price
```{r}
summarise(demo, sdcarpr=sd(carpr))
```
# select from the demo2 data the owners that are older than 40
```{r}
demo3 <- filter(demo2, age>40)
View(demo3)
```
# compute the average income in the demo3 data frame
```{r}
summarise(demo3, avginc=mean(income))
```
# compute the means  of columns 1 and 4 with the across function
```{r}
summarise(demo2, across(c(1,4), mean))
```
# compute the means of all numeric variables for the luxury cars only
```{r}
summarise(filter(demo2, carcat == 'Luxury'), across(where(is.numeric), mean))
```

# Count()
# count the rows in the demo data frame
```{r}
count(demo)
```
# count the rows where the income is smaller than 100
```{r}
count(filter(demo, income < 100))
```
# count the number of rows for each education level
```{r}
count(demo, educ)
```
# count the number of owners in each combination of  the variables gender and retired
```{r}
count(demo, gender, retired)
```
# the add_count function adds a column (variable) containing the number of rows
```{r}
demo2 <- add_count(demo)
View(demo2)
```

#tally()
# compute the income sum
```{r}
tally(demo, income)
```
# compute the car prices sum
```{r}
tally(demo, carpr)
```
# add a variable containing the number of rows
```{r}
demo2 <- add_tally(demo)
```
# add a variable containing the sum of the incomes
```{r}
demo3 <- add_tally(demo, income)
View(demo3)
```

# n_distinct()
# extract the variable educ as a vector and compute the number of unique values in this vector
```{r}
educ <- demo$educ
n_distinct(educ)
```
# extract the variable income as a vector and compute the number of unique values in this vector
```{r}
inc <- demo$income
n_distinct(inc)
```
# if the argument of the n_distinct() function is a data frame the function returns the number of rows
```{r}
n_distinct(demo)
```

# sample()
# the sample_n() function extracts a number of rows, at random
```{r}
demo_s <- sample_n(demo, 100)
View(demo_s)
```
# sampling with replacement
```{r}
demo_s <- sample_n(demo, 100, replace = TRUE)
View(demo_s)
```
# when we do sampling with replacement, the number of extracted rows can be greater than the number of rows in the initial data frame
# create a data frame with 5 rows and extract 100 rows with replacement
```{r}
id <- 1:5
names <- c("Tom", "Mary", "Nicolas", "Diana", "Ellen")
dat <- data_frame(id, names)
View(dat)
dat_s <- sample_n(dat, 100, replace = TRUE)
View(dat_s)
```
# sample_frac() extracts a fraction of rows at random
# extract 30% of rows
```{r}
demo_s2 <- sample_frac(demo, 0.3)
View(demo_s)
```
# extract 30% of rows with replacement
```{r}
demo_s2 <- sample_frac(demo, 0.3, replace = TRUE)
View(demo_s2)
```
# top_n()
# extract the top 10 owners based on income
```{r}
demo2 <- top_n(demo, 10, income)
View(demo2)
```
# extract the top 10 owners based on age
```{r}
demo2 <- top_n(demo, 10, age)
View(demo2)
```
# bind()
# create a data frame that contains the first 10 rows of demo
```{r}
demo1 <- demo[1:10,]
View(demo1)
```
# create a data frame that contains the last 10 row od demo
```{r}
demo2 <- demo[501:510,]
View(demo2)
```
# bind the two data frame with bind_row()
```{r}
demo_b <- bind_rows(demo1, demo2)
View(demo_b)
```
# bind_cols()
# create a data frame that contains the first 3 columns of demo
```{r}
demo1 <- demo[,1:3]
View(demo1)
```
# create a data frame that contains the last 5 columns of demo
```{r}
demo2 <- demo[,4:8]
View(demo2)
```
# bind the two data frames by columns
```{r}
demo_b <- bind_cols(demo1, demo2)
View(demo_b)
```
# glimpse
# the glimpse() function returns a summary of our data frame
```{r}
glimpse(demo)
```

# set operations
# intersect
```{r}
x <- 1:10
y <- c(2,4,6,9,11,15)
intersect(x, y)
```
# create a data frame that contains the rows 1 to 30 of the demo data set
```{r}
demo1 <- demo[1:30,]
View(demo1)
```
# create a data frame that contains rows 20 to 40
```{r}
demo2 <- demo[20:40,]
View(demo2)
```
# intersect the data frames
```{r}
intersect(demo1, demo2)
```
# union
```{r}
x <- 1:10
y <- c(2,4,6,9,11,15)
union(x, y)
```
```{r}
union(demo1, demo2)
```
# setdiff()
# get the elements that are found in x, but not in y
```{r}
x <- 1:10
y <- c(2,4,6,9,11,15)
setdiff(x, y)
```
# get the elements that are found in y, but not x
```{r}
setdiff(y, x)
```
# find uncommon data frame rows
```{r}
demo1 <- demo[1:30,]
View(demo1)

demo2 <- demo[20:40,]
View(demo2)

setdiff(demo1, demo2)
setdiff(demo2, demo1)
```
# setequal()
```{r}
x <- 1:10
y <- c(2,4,6,9,11,15)
setequal(x, y)
```
```{r}
w <- c(20, 30, 40)
z <- c(20, 30, 40)
setequal(w, z)
```
```{r}
setequal(demo1, demo2)
```

# chaining in dplyr
```{r}
# select the owners younger than 40 and with some college education
demo2 <- demo %>% filter(age < 40 & educ == "some college")
View(demo2)
# retain the income, car category and car price variables
demo2 <- demo %>% select(income, carcat, carpr)
View(demo2)
# select the owners younger than 40 and with some college education then retain the variables income, car category and car price only
demo2 <- demo %>% filter(age < 40 & educ == "Some college") %>% select(income, carcat, carpr)
View(demo2)
# select the owners younger than 40 and with some college education, retain the variables income, car category and car price only then compute a new variable "x" as income divided to price
demo2 <- demo %>% filter(age < 40 & educ == "Some college") %>% select(income, carcat, carpr) %>% mutate(x = income/carpr)
View(demo2)
# select owners younger than 40 and with some college education, retain the variables income, car category and car price only, then sort the data frame by income in descending order
demo2 <- demo %>% filter(age < 40 & educ == "Some college") %>% select(income, carcat, carpr) %>% arrange(desc(income))
View(demo2)
# select the owners younger than 40 and with some college education, then compute the average income
demo %>% filter(age < 40 & educ == "Some college") %>% summarise(avginc=mean(income, na.rm = TRUE))
# select the owners younger than 40 and with some college education, then compute the mean and the standard deviation of income
demo %>% filter(age < 40 & educ == "Some college") %>% summarise(avginc=mean(income, na.rm = TRUE), sdinc=sd(income, na.rm = TRUE))
# compute the means of all the numeric variables in the demo data frame for each gender category separately
demo %>% group_by(gender) %>% summarise(across(where(is.numeric), mean))
# select the owners of luxury cars and compute the means of the numeric variables by gender category
demo %>% group_by(gender) %>% filter(carcat == "Luxury") %>%  summarise(across(where(is.numeric), mean))
# select the owners of luxury cars, then compute the maximum values of the numeric variables by gender category
demo %>% group_by(gender) %>% filter(carcat == "Luxury") %>% summarise(across(where(is.numeric), max))
# select the owners of luxury cars, then return the number of owners in each gender category
demo %>% group_by(gender) %>% filter(carcat == "Luxury") %>% summarise(n())
# retain the income and car price only, then compute the means of these variables in each gender category
demo %>% group_by(gender) %>% select(carpr, income) %>% summarise(across(everything(), mean))
# retain the variables income and car price, then compute the variable "x" as income divided to price, then sort the data frame by the new variable x, in ascending order
demo2 <- demo %>% select(carpr, income) %>% mutate(x=income/carpr) %>% arrange(x)
View(demo2)
# select the owners with the income over 40 and return their number for each marital status- gender combination
demo %>% filter(income > 40) %>% count(marital, gender)
# select the owners with the age over 40 and compute the sum of their income
demo %>% filter(age > 40) %>% tally(income)
# select the owners with the income over 40 and then compute the sum of their incomes by marital status
demo %>% group_by(marital)  %>% filter(income > 40) %>% tally(income)
# select the owners of standard cars and extract a sample of 50 owners at random
demo2 <- demo %>% filter(carcat == "Standard") %>% sample_n(50)
View(demo2)
# select the owners of standard cars, retain the variables age, income, car price and car category, then extract a sample of 50 owners at random
demo2 <- demo %>% filter(carcat == "Standard") %>% select(age, income, carpr, carcat) %>% sample_n(50)
View(demo2)
# select the owners of standard cars, retain the variables age, income, car price, and car category, then extract a sample of 50 owners at random, then get the top 5 of them by income
demo2 <- demo %>% filter(carcat == "Standard") %>% select(age, income, carpr, carcat) %>% sample_n(50) %>% top_n(5, income)
View(demo2)
# select the owners of standard cars, retain the variables age, income, car price, and car category, then extract a sample of 50 owners at random, then get a glimpse of the new data frame demo2
demo2 <- demo %>% filter(carcat == "Standard") %>% select(age, income, carpr, carcat) %>% sample_n(50) %>% glimpse
```

# joining data frames
# inner_join()
# inner join returns all rows from the first data frame where there are matching values in the second data frame, and all variables from both data frames
```{r}
cities <- read.csv("cities.csv")
View(cities)
buildings <- read.csv("buildings.csv")
View(buildings)

# inner join cities with building
ij <- inner_join(cities, buildings)
View(ij)
```
# semi_join()
# this command returns all rows from the first data frame where there are matching values in the second data frame, keeping variables from the first data frame only
```{r}
sj <- semi_join(cities, buildings)
View(sj)

sj2 <- semi_join(buildings, cities)
View(sj2)
```
# left_join()
# this command returns all rows from the first data frame and all variables from both data frames
```{r}
lj <- left_join(cities, buildings)
View(lj)
```
# anti_join()
# returns all rows from the first data frame where there are NOT matching values in the second data frame and retains the variables from the first data frame
```{r}
aj <- anti_join(cities, buildings)
View(aj)

aj2 <- anti_join(buildings, cities)
View(aj2)
```
# full_join
# returns all rows and variables from both data frames, where there are no matches, the NA values is returned
```{r}
fj <- full_join(cities, buildings)
View(fj)

fj2 <- full_join(buildings, cities)
View(fj2)
```

# dplyr and ggplot2
# column chart
```{r}
require(ggplot2)
# create a column chart that represents the average car price by car category
demo %>% group_by(carcat) %>%
  summarise(avgpr=mean(carpr, na.rm = TRUE)) %>%
  ggplot(aes(x=carcat,y=avgpr))+
  geom_col(fill="darkblue")

# a column chart that displays the maximum car price by car category
demo %>% group_by(carcat) %>%
  summarise(maxpr=max(carpr, na.rm = TRUE)) %>%
  ggplot(aes(x=carcat, y=maxpr))+
  geom_col(fill="green")

# a column chart that represents the average income by gender for the luxury car owners
demo %>% group_by(gender) %>%
  filter(carcat=="Luxury") %>%
  summarise(avginc=mean(income, na.rm = TRUE)) %>%
  ggplot(aes(x=gender, y=avginc))+
  geom_col(fill="blue")
```
# mean plot chart
```{r}
# a mean  plot chart that displays the average income by educational level
demo %>% group_by(educ) %>%
  summarise(avginc=mean(income, na.rm = TRUE)) %>%
  ggplot(aes(x=educ, y=avginc, group=1))+
  geom_line(color="darkred")

# a mean plot chart that displays the average car price by gender
demo %>% group_by(gender) %>%
  summarise(avgpr=mean(carpr, na.rm = TRUE)) %>%
  ggplot(aes(x=gender, y=avgpr, group=1))+
  geom_line(color="darkred")

# a mean plot chart that represents the average car price by gender for the owners with the income greater than 70
demo %>% group_by(gender) %>%
  filter(income>70) %>%
  summarise(avgpr=mean(carpr, na.rm = TRUE)) %>%
  ggplot(aes(x=gender, y=avgpr, group=1))+
  geom_line(color="darkred")

# a mean plot chart that represents the average income by gender with a separate line for each marital status
demo %>% group_by(gender, marital) %>%
  summarise(avginc=mean(income, rm.na = TRUE)) %>%
  ggplot(aes(x=gender, y=avginc, group=marital,
             color=marital))+geom_line()

# a mean plot chart that represents the average income by marital status with a separate line for each gender
demo %>% group_by(marital, gender) %>%
  summarise(avginc=mean(income, na.rm = TRUE)) %>%
  ggplot(aes(x=marital, y=avginc, group=gender, color=gender))+geom_line()

# average income by marital status, with a separate line for each gender, for economy car owners only
demo %>% group_by(marital, gender) %>%
  filter(carcat=="Economy") %>%
  summarise(avginc=mean(income, na.rm = TRUE)) %>%
  ggplot(aes(x=marital, y=avginc, group=gender, color=gender))+geom_line()
```
# scatterplot
```{r}
mk <- read.csv("marketingdb.csv")
View(mk)

# a scatterplot that respresents the relationship between age and income by gender
mk %>% ggplot(aes(x=age, y=income, color=gender))+
  geom_point()

# a scatterplot that represents the relationship between age and income by gender, for the Mastercard owners only
mk %>% filter(card == "Mastercard") %>%
  ggplot(aes(x=age, y=income, color=gender))+
  geom_point()

# a scatterplot that represents the relationship between age and income by gender, for the people with at least 20 years of education
mk %>% filter(ed >= 20) %>%
  ggplot(aes(x=age, y=income, color=gender))+
  geom_point()
```
# histograms
```{r}
# histogram of the income by gender category using the demo dataset
demo %>% 
  ggplot(aes(x=income, fill=gender))+
  geom_histogram()

# histogram of income by gender category for the luxury car owners
demo %>%
  filter(carcat == "Luxury") %>%
  ggplot(aes(x=income, fill=gender))+
  geom_histogram()

# histogram of income by gender category for the car prices lower than 50
demo %>%
  filter(carpr < 50) %>%
  ggplot(aes(x=income, fill=gender))+
  geom_histogram()
```
# boxplot
```{r}
# boxplot of income by gender category
demo %>%
  ggplot(aes(x=gender, y=income))+
  geom_boxplot()

# boxplot of income by gender category for the standard car owners
demo %>%
  filter(carcat == "Standard") %>%
  ggplot(aes(x=gender, y=income))+
  geom_boxplot()

# boxplot of income by gender category for the owners aged between 35 and 50
demo%>%
  filter(age>=35 & age <=50) %>%
  ggplot(aes(x=gender, y=income))+
  geom_boxplot()
```

# perform the following operations using the data frame incomedata.csv
```{r}
inc <- read.csv("incomedata.csv")
View(inc)
# select the variables age category (agecat) and gender for the subjects who did not complete high school
inc2 <- inc %>% filter(educ == "Did not complete high school") %>% select(agecat, gender)
View(inc2)
# select the variables age category and gender for the subjects who did not complete high school and have the income greater than 20
inc2 <- inc %>% filter(educ == "Did not complete high school" & income > 20) %>% select(agecat, gender)
View(inc2)
# select the male subjects with the income lower than 65 and compute their average income
inc2 <- inc %>% filter(gender == "Male" & income < 65) %>% summarise(avginc=mean(income, na.rm = TRUE))
View(inc2)
# group subjects age category and compute the mean and the standard deviation of the income
inc2 <- inc %>% group_by(agecat) %>% summarise(avginc=mean(income, na.rm = TRUE), sdinc=sd(income, na.rm = TRUE))
View(inc2)
# sort the data frame by income, in ascending and descending order 
inc2 <- inc %>% arrange(desc(income))
View(inc2)
inc2 <- inc %>% arrange(income)
View(inc2)
# count the number of subjects in each education category
inc2 <- inc %>% group_by(educ) %>% summarise(n())
View(inc2)
```